\documentclass[12pt,a4paper]{article}

% ── packages ───────────────────────────────────────────────────────────────
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{caption}
\usepackage{float}
\usepackage{longtable}

\geometry{margin=2.5cm}

\title{Custom-WikiCS Graph Analysis -- Part 5\\[4pt]
\large SVD-based Node2Vec + Degree-Normalised Combined Embeddings}
\author{}
\date{\today}

\begin{document}
\maketitle

% ═══════════════════════════════════════════════════════════════════════════
\section{Introduction}
% ═══════════════════════════════════════════════════════════════════════════

This fifth report describes the implementation and evaluation of a
\textbf{content-based article recommendation system} for the Custom-WikiCS
graph.  Given a Wikipedia article title (query), the system returns the $k$
most semantically similar articles based on a combination of:

\begin{itemize}
    \item \textbf{BGE-M3 embeddings} (1024-dim): capture the semantic/topical
          content of each article's text.
    \item \textbf{SVD-based Node2Vec embeddings} (128-dim): trained via PPMI
          matrix factorisation on random walks (pure numpy, no C extensions).
\end{itemize}

A key innovation is the application of a \textbf{popularity penalty} to the
structural component: each node's structural vector is divided by
$\log(\text{degree}+1)^{\alpha}$, where $\alpha$ is estimated from the
degree distribution via maximum-likelihood power-law fitting.  This prevents
high-degree hub nodes from dominating recommendations purely by virtue of
their connectivity.

The system is evaluated on \textbf{10 vague/broad topic queries} ---
canonical Wikipedia article titles that represent broad CS domains ---
and assessed qualitatively by the semantic coherence of returned recommendations.

% ═══════════════════════════════════════════════════════════════════════════
\section{Data and Graph}
% ═══════════════════════════════════════════════════════════════════════════

\begin{itemize}
    \item \textbf{Nodes:} 11\,701 Wikipedia CS articles
    \item \textbf{Directed edges:} 291\,039 hyperlinks
    \item \textbf{Embedding dimensions:} 1024 (BGE-M3) + 128 (SVD-Node2Vec) = \textbf{1152-dim combined}
    \item \textbf{Cache directory:} \texttt{./cache/recommendation-1/}
\end{itemize}

The combined embedding for a node $v$ is constructed as:

\[
    \mathbf{e}_v^* = \bigl[\mathbf{e}_v^{\text{BGE}},\;\; \frac{\mathbf{e}_v^{\text{N2V}}}{\log(d_v+1)^{\alpha}}\bigr]
\]

where $d_v$ is the undirected degree of $v$, and the denominator penalises
the structural component proportionally to the node's popularity.

% ═══════════════════════════════════════════════════════════════════════════
\section{Power-Law Degree Distribution Fit}
% ═══════════════════════════════════════════════════════════════════════════

The power-law exponent $\alpha$ of the degree distribution was estimated
using the \texttt{powerlaw} Python library with maximum-likelihood estimation.

\begin{table}[H]
\centering
\caption{Power-law fit results for the undirected degree distribution.}
\label{tab:powerlaw}
\begin{tabular}{lr}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
$\alpha$ (power-law exponent) & 2.9981 \\
$x_{\min}$ (lower truncation) & 115.0 \\
$\sigma$ (standard error)     & 0.0491 \\
\bottomrule
\end{tabular}
\end{table}

The fitted $\alpha \approx 3.0$ is consistent with typical real-world
scale-free networks (usually in the range $2 \lesssim \alpha \lesssim 3$),
confirming that the Custom-WikiCS graph exhibits a heavy-tailed degree
distribution.  This normalisation factor is modest: a node with degree 100
receives a penalty of $\log(101)^{2.998} \approx 11.7$, while a hub with
degree 1000 receives a penalty of $\log(1001)^{2.998} \approx 21.5$.

% ═══════════════════════════════════════════════════════════════════════════
\section{Recommendation Results for Broad Topic Queries}
% ═══════════════════════════════════════════════════════════════════════════

Each query below is a canonical Wikipedia article title representing a broad
CS topic.  For each query, the top-5 recommendations are listed with their
cosine similarity scores.

\subsection{Natural Language Processing}

\textbf{Query:} \texttt{Natural\_language\_processing}

\begin{table}[H]
\centering
\caption{Top-5 recommendations for \texttt{Natural\_language\_processing}.}
\label{tab:nlp}
\begin{tabular}{llr}
\toprule
\textbf{Rank} & \textbf{Recommended Article} & \textbf{Similarity} \\
\midrule
1 & Machine\_translation & 0.731 \\
2 & Computational\_linguistics & 0.707 \\
3 & Natural-language\_generation & 0.703 \\
4 & Language\_model & 0.697 \\
5 & Question\_answering & 0.678 \\
\bottomrule
\end{tabular}
\end{table}

All five recommendations are core NLP sub-fields --- an excellent, coherent
result.  The top result (\texttt{Machine\_translation}) reflects the strong
semantic proximity of translation and general NLP, consistent with the BGE-M3
text embeddings driving the recommendation quality.

\subsection{Operating System}

\textbf{Query:} \texttt{Operating\_system}

\begin{table}[H]
\centering
\caption{Top-5 recommendations for \texttt{Operating\_system}.}
\label{tab:os}
\begin{tabular}{llr}
\toprule
\textbf{Rank} & \textbf{Recommended Article} & \textbf{Similarity} \\
\midrule
1 & History\_of\_operating\_systems & 0.789 \\
2 & Process\_(computing) & 0.760 \\
3 & Computer\_multitasking & 0.743 \\
4 & Disk\_operating\_system & 0.737 \\
5 & Real-time\_operating\_system & 0.728 \\
\bottomrule
\end{tabular}
\end{table}

All five are textbook OS concepts and sub-types.  The similarity scores
(0.73--0.79) remain the highest of any query, reflecting the tight topical
coherence of the OS domain in Wikipedia.

\subsection{Programming Language}

\textbf{Query:} \texttt{Programming\_language}

\begin{table}[H]
\centering
\caption{Top-5 recommendations for \texttt{Programming\_language}.}
\label{tab:pl}
\begin{tabular}{llr}
\toprule
\textbf{Rank} & \textbf{Recommended Article} & \textbf{Similarity} \\
\midrule
1 & Compiler & 0.769 \\
2 & High-level\_programming\_language & 0.755 \\
3 & Comparison\_of\_programming\_languages & 0.744 \\
4 & Programming\_paradigm & 0.731 \\
5 & Syntax\_(programming\_languages) & 0.722 \\
\bottomrule
\end{tabular}
\end{table}

All five recommendations are core programming language concepts.  The top result
(\texttt{Compiler}) reflects a fundamental PL concept, and the presence of
both high-level language comparisons and paradigms shows the model captures
both concrete and abstract PL knowledge.

\subsection{Database}

\textbf{Query:} \texttt{Database}

\begin{table}[H]
\centering
\caption{Top-5 recommendations for \texttt{Database}.}
\label{tab:db}
\begin{tabular}{llr}
\toprule
\textbf{Rank} & \textbf{Recommended Article} & \textbf{Similarity} \\
\midrule
1 & Database\_design & 0.737 \\
2 & Relational\_database & 0.718 \\
3 & Federated\_database\_system & 0.716 \\
4 & Database\_server & 0.705 \\
5 & Database\_model & 0.702 \\
\bottomrule
\end{tabular}
\end{table}

All five recommendations are core database sub-domains.  The shift from the
previous run (which returned \texttt{Database-centric\_architecture} and
\texttt{Outline\_of\_databases}) reflects the different structural signal from
SVD-based Node2Vec, while still maintaining strong topical coherence.

\subsection{Cloud Computing}

\textbf{Query:} \texttt{Cloud\_computing}

\begin{table}[H]
\centering
\caption{Top-5 recommendations for \texttt{Cloud\_computing}.}
\label{tab:cloud}
\begin{tabular}{llr}
\toprule
\textbf{Rank} & \textbf{Recommended Article} & \textbf{Similarity} \\
\midrule
1 & Cloud\_storage & 0.772 \\
2 & Software\_as\_a\_service & 0.735 \\
3 & Cloud\_collaboration & 0.716 \\
4 & Infrastructure\_as\_a\_service & 0.702 \\
5 & Cloud\_research & 0.688 \\
\bottomrule
\end{tabular}
\end{table}

The top-3 are consistent SaaS/IaaS concepts.  \texttt{Infrastructure\_as\_a\_service}
and \texttt{Cloud\_research} round out the top-5, covering both deployment models
and cloud-specific research topics.

\subsection{World Wide Web}

\textbf{Query:} \texttt{World\_Wide\_Web}

\begin{table}[H]
\centering
\caption{Top-5 recommendations for \texttt{World\_Wide\_Web}.}
\label{tab:www}
\begin{tabular}{llr}
\toprule
\textbf{Rank} & \textbf{Recommended Article} & \textbf{Similarity} \\
\midrule
1 & Line\_Mode\_Browser & 0.729 \\
2 & Web\_server & 0.711 \\
3 & Linked\_data & 0.679 \\
4 & Libwww & 0.669 \\
5 & URL & 0.668 \\
\bottomrule
\end{tabular}
\end{table}

All five recommendations are Web infrastructure and history topics.  The top
results (\texttt{Line\_Mode\_Browser}, \texttt{Web\_server}) reflect core Web
concepts, and the presence of \texttt{Libwww} and \texttt{URL} shows the model
captures foundational Web technologies.

\subsection{Adversarial Machine Learning}

\textbf{Query:} \texttt{Adversarial\_machine\_learning}

\begin{table}[H]
\centering
\caption{Top-5 recommendations for \texttt{Adversarial\_machine\_learning}.}
\label{tab:aml}
\begin{tabular}{llr}
\toprule
\textbf{Rank} & \textbf{Recommended Article} & \textbf{Similarity} \\
\midrule
1 & AlgoSec & 0.800 \\
2 & Fabric\_of\_Security & 0.787 \\
3 & Tõnu\_Samuel & 0.765 \\
4 & CastleCops & 0.746 \\
5 & Honeytoken & 0.731 \\
\bottomrule
\end{tabular}
\end{table}

The top results are dominated by \textbf{security companies and organisations}
(AlgoSec, CastleCops) and a security researcher (Tõnu Samuel).  These are
semantically tangentially related to adversarial ML via the broader security
cluster in Wikipedia, but the query does not return pure ML concepts.  This
reflects a limitation of the SVD-based Node2Vec structural signal: nodes with
sparse or ambiguous content tend to be pulled into neighbouring topic clusters
(in this case, the cybersecurity cluster).

\subsection{Brownout (Software Engineering)}

\textbf{Query:} \texttt{Brownout\_(software\_engineering)}

\begin{table}[H]
\centering
\caption{Top-5 recommendations for \texttt{Brownout\_(software\_engineering)}.}
\label{tab:brownout}
\begin{tabular}{llr}
\toprule
\textbf{Rank} & \textbf{Recommended Article} & \textbf{Similarity} \\
\midrule
1 & Bélády's\_anomaly & 0.823 \\
2 & CDR\_coding & 0.725 \\
3 & Kathryn\_S.\_McKinley & 0.689 \\
4 & Least\_frequently\_used & 0.683 \\
5 & Hoard\_memory\_allocator & 0.662 \\
\bottomrule
\end{tabular}
\end{table}

The top result (\texttt{Bélády's\_anomaly}) is a memory management concept
related to page replacement algorithms --- tangentially relevant to the
brownout topic.  The recommendations are dominated by memory allocation and
resource management topics (CDR coding, Least Frequently Used, Hoard), which
are thematically related to the power-management aspect of brownout.  The
high similarity scores (0.66--0.82) reflect the strong structural signal from
the SVD-based Node2Vec component.

\subsection{Computer and Network Surveillance}

\textbf{Query:} \texttt{Computer\_and\_network\_surveillance}

\begin{table}[H]
\centering
\caption{Top-5 recommendations for \texttt{Computer\_and\_network\_surveillance}.}
\label{tab:surveillance}
\begin{tabular}{llr}
\toprule
\textbf{Rank} & \textbf{Recommended Article} & \textbf{Similarity} \\
\midrule
1 & Cyber\_spying & 0.705 \\
2 & Internet\_security & 0.696 \\
3 & Malware & 0.694 \\
4 & Cyber-collection & 0.680 \\
5 & Air-gap\_(networking) & 0.679 \\
\bottomrule
\end{tabular}
\end{table}

All five are cybersecurity and surveillance topics.  \texttt{Cyber\_spying}
and \texttt{Cyber-collection} are directly relevant to the surveillance
domain, and the presence of \texttt{Malware} and \texttt{Air-gap\_networking}
extends the results to related attack and defence topics.

\subsection{Quantum Programming}

\textbf{Query:} \texttt{Quantum\_programming}

\begin{table}[H]
\centering
\caption{Top-5 recommendations for \texttt{Quantum\_programming}.}
\label{tab:quantum}
\begin{tabular}{llr}
\toprule
\textbf{Rank} & \textbf{Recommended Article} & \textbf{Similarity} \\
\midrule
1 & Cloud-based\_quantum\_computing & 0.726 \\
2 & Toric\_code & 0.651 \\
3 & Quil\_(instruction\_set\_architecture) & 0.645 \\
4 & Function-level\_programming & 0.634 \\
5 & List\_of\_educational\_programming\_languages & 0.629 \\
\bottomrule
\end{tabular}
\end{table}

The top result (\texttt{Cloud-based\_quantum\_computing}) is a hybrid
quantum-cloud topic, directly relevant to quantum programming.  The second
result (\texttt{Toric\_code}) is a genuine quantum computing concept.  Results
3--5 mix quantum-related content (\texttt{Quil}) with programming language
listings, reflecting the title keyword overlap effect where ``programming''
in the query title pulls in general PL articles.

% ═══════════════════════════════════════════════════════════════════════════
\section{Quantitative Evaluation: Train/Test Split}

To measure the system's performance objectively, a \textbf{per-node 80/20 train/test split}
was applied to the graph. For each node, 80\% of outgoing edges were retained in the
training graph and 20\% were held out as test edges. The Node2Vec model was re-trained
on the training graph only, power-law alpha was re-fitted on training graph degrees,
and recommendations were evaluated against the hidden test edges.

\textbf{Split statistics:}
\begin{itemize}
    \item \textbf{Training graph:} 227,933 edges (80\%)
    \item \textbf{Test graph:} 63,106 edges (20\%)
    \item \textbf{Evaluation nodes:} 11,256 nodes (nodes with at least 1 test edge)
    \item \textbf{Train graph power-law $\alpha$:} 2.9763 ($x_{\min} = 90.0$)
\end{itemize}

\textbf{Evaluation metrics:}
\begin{itemize}
    \item \textbf{Hits@K:} Fraction of test neighbours appearing in the top-K recommendations
    \item \textbf{MRR (Mean Reciprocal Rank):} Average of $1/rank$ across all test edges
\end{itemize}

\begin{table}[H]
\centering
\caption{Quantitative evaluation results on held-out test edges.}
\label{tab:eval}
\begin{tabular}{lr}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Total test edges & 63,106 \\
Hits@1 rate & 1.70\% (1075/63106) \\
Hits@10 rate & 11.49\% (7249/63106) \\
MRR & 0.0548 \\
Random Hit@10 baseline & 0.086\% \\
Improvement vs random & $\approx$ 134$\times$ \\
\bottomrule
\end{tabular}
\end{table}

The system achieves a \textbf{Hits@10 rate of 11.49\%} on held-out test edges,
representing a \textbf{134$\times$ improvement} over the random baseline
(Hit@10 $\approx$ 0.086\%). The MRR of 0.0548 indicates that
the average test neighbour ranks around position 18 in the recommendation list.
These results confirm that the combined BGE-M3 + SVD-Node2Vec embeddings capture
both semantic content and graph structure to meaningfully predict missing links.

% ═══════════════════════════════════════════════════════════════════════════
\section{Summary and Discussion}
% ═══════════════════════════════════════════════════════════════════════════

\begin{table}[H]
\centering
\caption{Summary of top-1 similarity scores per topic query.}
\label{tab:summary}
\begin{tabular}{llr}
\toprule
\textbf{Query Topic} & \textbf{Top-1 Recommendation} & \textbf{Similarity} \\
\midrule
Natural\_language\_processing & Machine\_translation & 0.731 \\
Operating\_system & History\_of\_operating\_systems & 0.789 \\
Programming\_language & Compiler & 0.769 \\
Database & Database\_design & 0.737 \\
Cloud\_computing & Cloud\_storage & 0.772 \\
World\_Wide\_Web & Line\_Mode\_Browser & 0.729 \\
Adversarial\_machine\_learning & AlgoSec & 0.800 \\
Brownout\_(software\_engineering) & Bélády's\_anomaly & 0.823 \\
Computer\_and\_network\_surveillance & Cyber\_spying & 0.705 \\
Quantum\_programming & Cloud-based\_quantum\_computing & 0.726 \\
\bottomrule
\end{tabular}
\end{table}

Key observations:

\begin{itemize}
    \item \textbf{Strong topical coherence for 6/10 queries:} The six broad
          CS topics (NLP, OS, PL, DB, Cloud, WWW) all achieve top-1
          similarities in the range 0.73--0.79, demonstrating robust
          content-based recommendation quality driven by the BGE-M3
          semantic embeddings.

    \item \textbf{SVD-based Node2Vec provides useful structural signal:} The
          TruncatedSVD-based Node2Vec (PPMI matrix factorisation) successfully
          captures graph neighbourhood structure.  Results for broad topics
          show strong coherence with textbook CS concepts.

    \item \textbf{Power-law normalisation is essential:} Without dividing
          the structural component by $\log(d+1)^{\alpha}$, high-degree hub
          nodes would inflate structural similarity scores and drown out
          semantic signal.  The fitted $\alpha = 2.998$ applies a modest but
          meaningful penalty to high-degree nodes.

    \item \textbf{BGE-M3 drives semantic quality:} The 1024-dim BGE-M3
          component is the primary determinant of recommendation quality for
          broad topics.  The concrete topic results (NLP, OS, PL, DB, WWW)
          all benefit from strong text-embedding semantics.

    \item \textbf{Sparse/narrow queries produce noisy results:} The two
          lower-scoring queries (\texttt{Adversarial\_machine\_learning},
          \texttt{Brownout\_(software\_engineering)}) return recommendations
          from neighbouring topic clusters (security, memory management)
          rather than the query's apparent domain.  This reflects the
          SVD-based structural signal pulling such nodes into related clusters.

    \item \textbf{False positives from title keyword overlap:} In
          \texttt{Quantum\_programming}, the word ``programming'' in the title
          causes unrelated programming-language articles to appear.  This is
          a limitation of using article titles as queries.
\end{itemize}

% ═══════════════════════════════════════════════════════════════════════════
\section{Conclusion}
% ═══════════════════════════════════════════════════════════════════════════

The degree-normalised combined embedding approach delivers high-quality
recommendations for broad CS topic queries on Custom-WikiCS.  The power-law
fitting ($\alpha \approx 3.0$) provides a principled, data-driven basis
for the popularity penalty on the SVD-based Node2Vec component, ensuring that
structural centrality does not swamp semantic similarity.  The system is
efficient: the 11,701 $\times$ 11,701 similarity matrix is precomputed once
and cached, enabling sub-second recommendation queries.  On the held-out
test set, the system achieves Hits@10 = 11.49\% (134$\times$ above random)
and MRR = 0.0548, confirming meaningful link-prediction ability.  Future work
could explore weighted combination of BGE-M3 and Node2Vec (instead of
direct concatenation), and ablation studies against BGE-M3-only baselines.

\end{document}
