\chapter{INTRODUCTION}
\thispagestyle{empty}

\section{Project Scope}

This project studies Wikipedia article recommendation in the computer science
domain. The long-term objective is not only to retrieve related articles, but
also to organize them into useful learning paths for a reader. The project
therefore combines three families of components:
\begin{itemize}
    \item semantic text embeddings obtained from article content,
    \item structural graph embeddings learned from Wikipedia hyperlink
    structure,
    \item graph neural network baselines and a future learning-path layer for
    presentation.
\end{itemize}

The experimental backbone of the work remains the Custom-WikiCS dataset, which
extends the Wiki-CS benchmark~\cite{mernyei2020wikics} with article text,
embedding caches, and evaluation artifacts tailored to recommendation-oriented
experiments. In the current branch, the project uses BGE-M3 text
embeddings~\cite{chen2025bgem3}, Node2Vec-style structural embeddings, a
Graph Convolutional Network (GCN) baseline~\cite{kipf2017semi}, and a
GraphSAGE baseline~\cite{hamilton2017inductive}. Graph Attention Networks
(GAT)~\cite{velickovic2018graph} remain part of the conceptual direction, but
the most stable results at this stage are obtained from the revised
evaluation branch centered on semantic-structural hybrid retrieval.

\section{Status at the Start of Mid-Report 3}

At the beginning of this third mid-report, the project has moved well beyond
problem definition and dataset construction. The following major milestones are
already complete:
\begin{itemize}
    \item construction of a cleaned Wikipedia article graph with 11,701 nodes
    and 291,039 directed edges,
    \item generation of BGE-M3 text embeddings and multiple Node2Vec-based
    structural branches,
    \item implementation of GCN and GraphSAGE structural baselines,
    \item development of a local web application prototype for interactive
    article retrieval,
    \item correction of earlier evaluation issues and creation of a revised
    undirected structural branch for fairer comparison.
\end{itemize}

The main purpose of this report is to present the first consolidated results of
that revised branch, analyze what changed compared with earlier experiments,
document the problems encountered during the process, and formalize the
performance metrics and evaluation protocol currently used for project
verification.
