\chapter{PERFORMANCE METRICS AND EVALUATION METHODS}
\thispagestyle{empty}

\section{Existing Evaluation Metrics}

For continuity with the earlier report series, the project still uses the
standard ranking metrics below:
\begin{itemize}
    \item \textbf{Hit@1, Hit@10, Hit@50, Hit@100:} proportion of test targets
    recovered within the first $K$ ranked positions,
    \item \textbf{MRR:} mean reciprocal rank of the true target over all
    evaluation queries,
    \item \textbf{NDCG:} normalized discounted cumulative gain, used to reflect
    the quality of the ranking position,
    \item \textbf{Recall:} fraction of relevant targets retrieved within the
    inspected cutoff,
    \item \textbf{MAP:} mean average precision, summarizing ranking quality over
    the evaluation set.
\end{itemize}

These metrics remain useful because they preserve comparability with earlier
branches and quantify how well the system recovers held-out graph relations.

\section{New Recommendation-Oriented Metrics}

Because the project is ultimately about recommendation rather than pure link
prediction, the revised branch adds a second family of metrics:
\begin{itemize}
    \item \textbf{Popularity Lift:} measures how strongly a model concentrates
    on popular articles compared with the baseline graph frequency,
    \item \textbf{Gini Index:} measures how concentrated or uneven the
    recommendation distribution is across the article set,
    \item \textbf{Aggregate Diversity:} counts how many distinct articles are
    surfaced across all recommendation lists,
    \item \textbf{Novelty:} rewards models that recommend less obvious articles,
    \item \textbf{Intra-List Distance (ILD):} measures how diverse the contents
    of each recommendation list are.
\end{itemize}

These metrics are important because two models with similar Hit@10 values can
still behave very differently as recommendation systems. A model may recover
many correct links while repeatedly surfacing the same high-degree or highly
popular nodes. The new metric family makes those tradeoffs visible.

\section{Evaluation Protocol}

The current official evaluation protocol is based on the revised undirected
structural branch. The procedure is summarized in
Table~\ref{tab:eval-branch-stats}.

\begin{table}[H]
\centering
\caption{Revised evaluation branch statistics and protocol parameters.}
\label{tab:eval-branch-stats}
\begin{tabular}{p{0.52\textwidth}p{0.28\textwidth}}
\toprule
\textbf{Quantity} & \textbf{Value} \\
\midrule
Full directed graph edges & 291,039 \\
Full undirected structural edges & 216,123 \\
Training edges (undirected branch) & 172,900 \\
Test edges (undirected branch) & 43,223 \\
Expanded evaluation pairs & 86,446 \\
Evaluation nodes & 9,022 \\
Structural power-law coefficient & $\alpha \approx 2.9966$ \\
Recommendation list length for new metrics & Top-20 \\
Number of models in the revised comparison & 10 \\
\bottomrule
\end{tabular}
\end{table}

Under this protocol, each undirected test edge $\{u,v\}$ is expanded into two
directional evaluation queries, $u \rightarrow v$ and $v \rightarrow u$, for the
ranking metrics. Recommendation-oriented metrics are then computed on the top-20
lists produced for each evaluation node. This dual evaluation design allows the
project to assess both graph-recovery quality and practical recommendation
behavior within the same experimental branch.
